---
title: "erhcv"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

[Homepage](https://spgadou.github.io)

### CRAN

[https://cran.r-project.org/web/packages/erhcv/index.html](https://cran.r-project.org/web/packages/erhcv/index.html)

### Short description

Assesses the statistical significance of clusters for a given dataset through bootstrapping and hypothesis testing of a given matrix of empirical Spearman's rho, based on the technique of S. Gaiser et al. (2010). 

### Tree structure convention

In this package, a *tree structure* consists of nested lists, which is very natural. To explain the concept, let us use the following custom data.tree object.
The associated *tree structure* is then constructed as follows.
```{r 1}
treeStructure <- list(list(list(7, 8), 6, 5, list(9, 10)), 2, 1, list(3, 4))
```

Note that one can transform this tree into a data.tree object using tree2plot.
```{r 2}
library(erhcv); library(nCopula)
tree2plot(treeStructure, structure = TRUE) # data.tree
```
As a general guideline, a tree structure needs to be a list. Then, one distinguishes nodes from leafs by object types. Indeed, nodes are lists, and leafs are integers. The construction is analogous to the structure construction of the package [nCopula](https://cran.r-project.org/web/packages/nCopula/index.html).

### Main functions

Function | Description
----------| -------------
VerifyTree | Main function of the package, used to *clean* raw hclust clustering
hclust2tree | Transforms a hclust object into a tree structure (under the convention of this package)
tree2plot | Illustration of a tree (or data.tree structure)

## Working example

### Packages

For a proper demonstration of the package, we use the package *nCopula* to sample hierarchical data.

```{r 3}
# install.packages(erhcv); install.packages(nCopula)
```

### Dataset sampling and clustering

```{r 4}
library(nCopula); library(erhcv)

## Build structure
structure <- GEO(0.5, 1:2, list(GAMMA(1/2, 3:4, NULL),
                                GEO(0.3, 5:6, list(GAMMA(1/3, 7:8, NULL),
                                                   GAMMA(1/3, 9:10, NULL)))))
                                                   
## Sample from the structure
set.seed(123)
U.. <- rCompCop(1000, structure)

## Compute Spearman correlation matrix
Spearman <- cor(U.., method = "sp")

## Cluster Spearman matrix
distance <- dist(Spearman, method = "maximum")
clustering <- hclust(distance, method = "average")

## Transform clustering into nested lists
tree <- erhcv::hclust2tree(clustering)
erhcv::tree2plot(tree, structure = TRUE) # data.tree object
```

We rapidly see that the obtained structure is far from the original one. We then use tools from *erhcv* to eliminate unnecessary nodes, based on our (subjective) simplification level *alpha*.

### Clustering validation

Here, we make use of *VerifyTree* to chop down nodes of the clustering we obtained earlier. 

```{r 5}
## Clean the tree
alpha <- 1 # Severe simplification parameter
cleanedTree <- erhcv::VerifyTree(U..,
                                 alpha = alpha,
                                 distance.method = "maximum",
                                 hclust.method = "average")$Tree

## Visualize output
erhcv::tree2plot(cleanedTree, structure = TRUE)
```

Bellow is the associated tree structure.

```{r 6}
erhcv::tree2plot(cleanedTree, structure = FALSE)
```
